### 2024-11-26 paper/paper.qmd
Upload starter files

---
title: "financial Well Being Analysis"
# subtitle: "My subtitle if needed"
author: 
  - Xinxiang Gao
thanks: "Code and data are available at: [https://github.com/xgao28/financial_well_being_analysis](https://github.com/xgao28/financial_well_being_analysis)."
date: today
date-format: long
abstract: "This should be an abstract. The report is still being finalized."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
data <- read_parquet("../data/02-analysis_data/NFWBS_analysis_data.parquet")
```

# Introduction

# Data

```{r}

```

# Model

# Results

# Discussion

# Appendix

## Data cleaning

## Surveys, sampling, and observational data

# References