### 2024-11-26 paper/paper.qmd
Updated readme, paper on acknowledgements

---
title: "financial Well Being Analysis"
# subtitle: "My subtitle if needed"
author: 
  - Xinxiang Gao
thanks: "Code and data are available at: [https://github.com/xgao28/financial_well_being_analysis](https://github.com/xgao28/financial_well_being_analysis)."
date: today
date-format: long
abstract: "This should be an abstract. The report is still being finalized."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(ggplot2)
data <- read_parquet("../data/02-analysis_data/NFWBS_analysis_data.parquet")
```

# Introduction

# Data

```{r}
#| echo: false
#| warning: false
#| message: false
library(skimr)
skim_summary <- data %>%
  select(-PUF_ID) %>%  # Remove 'PUF_ID' column
  select(where(is.numeric)) %>%  # Keep only numeric columns
  skim()

# Round the numeric columns of the skim summary
skim_summary %>%
  mutate_if(is.numeric, ~round(., 3)) %>%
  select(-complete_rate, -n_missing, -numeric.hist) %>% 
  print()
```
```{r}
#| echo: false
#| warning: false
#| message: false
data_cleaned <- data[, !names(data) %in% "puf_id"]
```

```{r}
#| echo: false
#| warning: false
#| message: false
data_long <- data_cleaned %>%
  pivot_longer(
    cols = c(FWBscore, FSscore, LMscore, KHscore),
    names_to = "ScoreType", values_to = "Value"
  )

# Plot using ggplot2
ggplot(data_long, aes(x = Value)) +
  geom_histogram(binwidth = 1, color = "black", fill = "skyblue", alpha = 0.7) +
  facet_wrap(~ScoreType, scales = "free", ncol = 2) +
  labs(
    title = "Histograms of Financial Scores",
    caption = "FWBscore: Financial Well-being Scale\nFSscore: Financial Skill Scale\nLMscore: Lusardi and Mitchell Financial Knowledge Scale\nKHscore: Knoll and Houts Financial Knowledge Scale",
    x = "Score Value",
    y = "Frequency"
  ) +
  theme_minimal()
```

The composition of finalwt remains on investigation.

```{r}
#| echo: false
#| warning: false
#| message: false
ggplot(data_cleaned, aes(x = finalwt)) +
  geom_density(color = "black", fill = "skyblue", alpha = 0.7) +
  labs(
    title = "Distribution of Final Weight (finalwt)",
    x = "Final Weight (finalwt)",
    y = "Frequency"
  ) +
  theme_minimal()
```

# Model

# Linear Regression Model

The linear regression model can be expressed mathematically as:

$$y_i = \beta_0 + \beta_1 \cdot \text{LMscore}_i + \beta_2 \cdot \text{KHscore}_i + \beta_3 \cdot \text{FWBscore}_i + \beta_4 \cdot \text{FSscore}_i + \epsilon_i, \quad \epsilon_i \sim \mathcal{N}(0, \sigma^{2})$$

### Model Components
- $y_i$: The dependent variable, $\text{finalwt}_i$, for observation $i$.
- $\beta_0$: The intercept term (constant in the model).
- $\beta_1, \beta_2, \beta_3, \beta_4$: The slopes (coefficients) of the respective predictor variables:
  - $\text{LMscore}_i$: Lusardi and Mitchell financial knowledge scale score.
  - $\text{KHscore}_i$: Knoll and Houts financial knowledge scale score.
  - $\text{FWBscore}_i$: Financial well-being scale score.
  - $\text{FSscore}_i$: Financial skill scale score.

```{r}
#| echo: false
#| warning: false
#| message: false
library(knitr)
# Build a linear regression model
# Assuming other columns in the dataset are predictors
model <- lm(finalwt ~ LMscore + KHscore + FWBscore + FSscore, data = data_cleaned)

# Display the summary of the model
model_summary <- summary(model)

# Create a data frame with key statistics
lm_results <- data.frame(
  Term = rownames(model_summary$coefficients),
  Estimate = model_summary$coefficients[, "Estimate"],
  Std_Error = model_summary$coefficients[, "Std. Error"],
  t_value = model_summary$coefficients[, "t value"],
  p_value = model_summary$coefficients[, "Pr(>|t|)"]
)

# Round values for better readability
lm_results[, -1] <- round(lm_results[, -1], 3)

# Display the table using knitr::kable
kable(
  lm_results,
  caption = "Summary of Linear Regression Model"
)
```

# Results

# Discussion

# Appendix

## Data cleaning

## Surveys, sampling, and observational data

## Acknowledgements

We would like to express our gratitude to the developers and contributors of R [@citeR] as well as several R packages that were essential for the analysis and visualization of the data in this report. The following R packages provided indispensable tools and functionalities:

- `tidyverse` [@tidyverse]: A collection of R packages designed for data science, including dplyr, ggplot2, readr, purrr, and others, which greatly facilitated data manipulation, analysis, and visualization.

- `ggplot2` [@ggplot2]: An implementation of the Grammar of Graphics, which allowed us to create complex and aesthetically pleasing visualizations with ease.

- `knitr` [@knitr]: This package enabled us to perform data demonstration with tables.

- `styler` [@styler]: This package is helpful for styling the code.

- `arrow` [@arrow]: This package provides a convenient and efficient way to work with parquet format.

We are grateful to the Consumer Financial Protection Bureau (CFPB) for conducting the Financial Well-Being Survey [@cfpb_fwb_survey] and making the data publicly available. Their dedication to collecting and sharing such vital information significantly contributes to research on financial well-being.

Finally, we would like to thank all those who contributed to the development and maintenance of the R programming language and its ecosystem, as well as the broader open-source community, whose efforts make such research possible.

# References