### 2024-12-03 paper/paper.qmd
Refined clean_data and complete introduction

---
title: "financial Well Being Analysis"
# subtitle: "My subtitle if needed"
author: 
  - Xinxiang Gao
thanks: "Code and data are available at: [https://github.com/xgao28/financial_well_being_analysis](https://github.com/xgao28/financial_well_being_analysis)."
date: today
date-format: long
abstract: "This should be an abstract. The report is still being finalized."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(ggplot2)
library(car)
library(GGally)
library(broom)
library(gridExtra)
library(knitr)
data <- read_parquet("../data/02-analysis_data/NFWBS_analysis_data.parquet")
```

```{r}
#| echo: false
#| warning: false
#| message: false
data_cleaned <- data[, !names(data) %in% "puf_id"]
data_cleaned <- data_cleaned %>%
  mutate(across(all_of(c("agecat", "PPEDUC", "PPINCIMP", "PPMARIT", "HEALTH", "DISTRESS")), as.factor))
```

# Introduction

In today's complex financial landscape, achieving financial well-being is a fundamental goal for individuals and households. While financial well-being encompasses the ability to meet current and future financial obligations, its determinants span multiple domains including financial literacy, demographic characteristics, and psychosocial factors. Understanding these relationships is crucial for developing effective policies and interventions to improve financial outcomes across diverse population segments.

To measure financial well-being systematically, the Consumer Financial Protection Bureau (CFPB) developed the Financial Well-Being Scale (FWBscore), a 10-item questionnaire assessing four critical elements: control over day-to-day finances, capacity to absorb financial shocks, progress toward financial goals, and ability to make life-improving choices. Alongside this measure, researchers have developed various instruments to assess financial knowledge and skills. The Lusardi and Mitchell financial knowledge scale (LMscore) @lusardi2008 provides a foundational three-item measure of basic financial concepts, while the Knoll and Houts financial knowledge scale (KHscore) @knoll2012 offers a more comprehensive 10-item assessment using item response theory. Additionally, the CFPB Financial Skill Scale (FSscore) @cfpb2015 evaluates practical financial competencies through an IRT-based approach.

While previous research has examined financial literacy's role in economic outcomes, less attention has been paid to how demographic and psychosocial factors interact with financial knowledge to influence overall financial well-being. This gap is particularly significant given evidence suggesting that financial knowledge alone may not fully explain variations in financial outcomes across different population segments.

This paper investigates how demographic characteristics (age, education, income, marital status) and psychosocial factors (health status, psychological distress) influence individuals' financial well-being as measured by the FWBscore, while accounting for their financial knowledge and skills as measured by the LMscore, KHscore, and FSscore. By examining these relationships comprehensively, we aim to identify which population segments may be more vulnerable to financial challenges and how personal circumstances interact with financial literacy to shape financial well-being outcomes.

Our findings indicate [brief summary of key results], suggesting important implications for financial education programs and policy interventions. These results highlight the need for tailored approaches that consider both financial knowledge and individual circumstances in promoting financial well-being.

The rest of the paper is organized as follows: Section 2 reviews relevant literature on financial well-being and its determinants. Section 3 describes our data and methodology. Section 4 presents our empirical results. Section 5 discusses the implications of our findings and concludes.

# Data

## Overview of the Dataset

The dataset for this study is the Consumer Financial Protection Bureau’s (CFPB) National Financial Well-Being Survey Public Use File (PUF), collected in 2017. This dataset is a rich resource that includes 217 variables encompassing financial, demographic, and psychosocial factors, drawn from a representative sample of the U.S. population. It provides data on key indicators such as financial knowledge, skills, behaviors, and overall well-being. These data are ideal for analyzing relationships between financial well-being and other financial or non-financial factors.

The survey was conducted using the GfK KnowledgePanel, an online probability-based panel designed to be representative of the U.S. adult population. The GfK KnowledgePanel is created by randomly selecting households and inviting them to participate in the panel. Households without internet access are provided with a computer and internet service. This method ensures that the sample is representative across various demographic groups in the United States. Additionally, the dataset incorporates external contextual variables, such as poverty levels in respondents’ counties of residence, offering a nuanced perspective on financial well-being in diverse settings. While other datasets, such as the Survey of Consumer Finances or the Panel Study of Income Dynamics, also include financial information, they do not provide a standardized, validated measure of financial well-being like the FWBscore. This makes the CFPB dataset uniquely suited for this study.

## Variables of Interest

### Primary Outcome Variable

Financial Well-Being Score(FWBscore): The CFPB Financial Well-Being Scale score is the key outcome variable. It is a 10-item scale capturing four dimensions: control over day-to-day finances, capacity to absorb financial shocks, progress toward financial goals, and freedom of financial choices. Scores range from 0 to 100, with higher scores indicating greater financial well-being.

### Independent Variables

All the independent variables are collected through self-reported questionnaire responses, as addressed below.

Age (agecat): Categorized into age groups (e.g., 18–24, 25–34, 35–44, 45–54, 55–64, 65+).

Education (PPEDUC): Highest degree attained, categorized from less than high school to graduate degrees.

Household Income (PPINCIMP): Categorized income levels (e.g., <$20,000, $20,000-$39,999, $40,000-$59,999, etc.).

Marital Status (PPMARIT): Categories include single, married, divorced, widowed, and other.

Health (HEALTH): Self-reported general health, rated on a 5-point Likert scale from 1 (Poor) to 5 (Excellent).

Stress (DISTRESS): Perceived level of stress, rated on a 5-point Likert scale from 1 (Not at all) to 5 (Extremely).

### Additional Variables

Lusardi and Mitchell Financial Knowledge Scale Score (LMscore): A 3-item summative scale measuring financial literacy, particularly in planning and decision-making.

Knoll and Houts Financial Knowledge Scale Score (KHscore): A 10-item IRT-based measure assessing broader financial knowledge across multiple dimensions.

Financial Skill Scale Score (FSscore): An IRT-based measure developed by the CFPB, capturing practical financial skills like budgeting, debt management, and saving.

```{r}
#| echo: false
#| warning: false
#| message: false
#|

# List of variables to plot
variables <- c("agecat", "PPEDUC", "PPINCIMP", "PPMARIT", "HEALTH", "DISTRESS")

# Initialize an empty list to store plots
plot_list <- list()

# Loop through variables and generate bar plots
for (var in variables) {
  plot <- ggplot(data_cleaned, aes_string(x = var)) +
    geom_bar(alpha = 0.7, color = "black", fill = "skyblue") +
    labs(
      title = paste("Distribution of", var),
      x = var,
      y = "Frequency"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      # axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    coord_flip()
  # Add the plot to the list
  plot_list[[var]] <- plot
}

grid.arrange(grobs = plot_list, ncol = 2)

```
```{r}
#| echo: false
#| warning: false
#| message: false
# Ensure categorical variables are factors
data_cleaned <- data_cleaned 
```

```{r}
#| echo: false
#| warning: false
#| message: false
data_long <- data_cleaned %>%
  pivot_longer(
    cols = c(FWBscore, FSscore, LMscore, KHscore),
    names_to = "ScoreType", values_to = "Value"
  )

ggplot(data_long, aes(x = Value)) +
  geom_histogram(binwidth = 1, color = "white", fill = "steelblue", alpha = 0.9) +
  facet_wrap(~ScoreType, scales = "free", ncol = 2) +
  labs(
    title = "Histograms of Financial Scores",
    subtitle = "Distribution of scores across different financial measures",
    caption = "FWBscore: Financial Well-being Scale\nFSscore: Financial Skill Scale\nLMscore: Lusardi and Mitchell Financial Knowledge Scale\nKHscore: Knoll and Houts Financial Knowledge Scale",
    x = "Score Value",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    strip.background = element_rect(fill = "lightgray"),
    strip.text = element_text(face = "bold")
  )
```

# Model

The linear regression model can be expressed mathematically as:

$$y_i = \beta_0 + \beta_1 \cdot \text{LMscore}_i + \beta_2 \cdot \text{KHscore}_i + \beta_3 \cdot \text{FSscore}_i + \epsilon_i, \quad \epsilon_i \sim \mathcal{N}(0, \sigma^{2})$$

### Model Components
- $y_i$: The dependent variable, $\text{FWBscore}_i$, for observation $i$.
- $\beta_0$: The intercept term (constant in the model).
- $\beta_1, \beta_2, \beta_3, \beta_4$: The slopes (coefficients) of the respective predictor variables:
  - $\text{LMscore}_i$: Lusardi and Mitchell financial knowledge scale score.
  - $\text{KHscore}_i$: Knoll and Houts financial knowledge scale score.
  - $\text{FWBscore}_i$: Financial well-being scale score.
  - $\text{FSscore}_i$: Financial skill scale score.

# Results

```{r}
#| echo: false
#| warning: false
#| message: false

# Build a linear regression model
# Assuming other columns in the dataset are predictors
model <- lm(FWBscore ~ LMscore + KHscore + FSscore, data = data_cleaned)

# Display the summary of the model
model_summary <- summary(model)

# Create a data frame with key statistics
lm_results <- data.frame(
  Term = rownames(model_summary$coefficients),
  Estimate = model_summary$coefficients[, "Estimate"],
  Std_Error = model_summary$coefficients[, "Std. Error"],
  t_value = model_summary$coefficients[, "t value"],
  p_value = model_summary$coefficients[, "Pr(>|t|)"]
)

# Round values for better readability
lm_results[, -1] <- round(lm_results[, -1], 3)

# Display the table using knitr::kable
kable(
  lm_results,
  caption = "Summary of Linear Regression Model"
)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| footnote: "The intercept represents the baseline category for each categorical variable. For instance, for 'agecat', the baseline is '18-24'; for 'PPEDUC', it is 'Less than high school'; for 'PPINCIMP', it is 'Less than $20,000'; for 'PPMARIT', it is 'Married'; for 'HEALTH', it is 'Poor'; and for 'DISTRESS', it is 'Strongly disagree'. All estimates are relative to these baselines."

# Preprocess data: Ensure predictors are factors and drop missing values
data_model <- data_cleaned %>%
  mutate(across(all_of(c("agecat", "PPEDUC", "PPINCIMP", "PPMARIT", "HEALTH", "DISTRESS")), as.factor))

# Fit the linear regression model
lm_model <- lm(FWBscore ~ agecat + PPEDUC + PPINCIMP + PPMARIT + HEALTH + DISTRESS, data = data_model)

# Create tidy output and filter for significant results
tidy_lm <- tidy(lm_model) %>% 
  mutate(Significant = ifelse(p.value < 0.05, "Yes", "No")) %>%
  filter(Significant == "Yes")

# Define variables and their mapping
variable_mapping <- tibble(
  term = tidy_lm$term,
  Variable = case_when(
    grepl("^agecat", term) ~ "agecat",
    grepl("^PPEDUC", term) ~ "PPEDUC",
    grepl("^PPINCIMP", term) ~ "PPINCIMP",
    grepl("^PPMARIT", term) ~ "PPMARIT",
    grepl("^HEALTH", term) ~ "HEALTH",
    grepl("^DISTRESS", term) ~ "DISTRESS",
    TRUE ~ "Intercept"
  ),
  Value = case_when(
    Variable == "Intercept" ~ "See footnote",
    TRUE ~ str_remove(term, Variable)
  )
)

# Join the mapping back to the original data
tidy_lm <- tidy_lm %>%
  left_join(variable_mapping, by = "term") %>%
  select(Variable, Value, estimate, std.error, statistic, p.value) %>%
  mutate(
    estimate = round(estimate, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = round(p.value, 3)
  )

# Create the table
library(knitr)
kable(tidy_lm, caption = "Linear Model Output with Significant Variable and Levels")
  # %>% footnote( "The intercept represents the baseline category for each categorical variable. For instance, for 'agecat', the baseline is '18-24'; for 'PPEDUC', it is 'Less than high school'; for 'PPINCIMP', it is 'Less than $20,000'; for 'PPMARIT', it is 'Married'; for 'HEALTH', it is 'Refused'; and for 'DISTRESS', it is 'Refused'. All estimates are relative to these baselines.")
```

```{r}
# Calculate VIF
library(car)
vif(lm_model)
```

# Discussion

# Appendix

## Data cleaning

## Surveys, sampling, and observational data

## Acknowledgements

We would like to express our gratitude to the developers and contributors of R [@citeR] as well as several R packages that were essential for the analysis and visualization of the data in this report. The following R packages provided indispensable tools and functionalities:

- `tidyverse` [@tidyverse]: A collection of R packages designed for data science, including dplyr, ggplot2, readr, purrr, and others, which greatly facilitated data manipulation, analysis, and visualization.

- `ggplot2` [@ggplot2]: An implementation of the Grammar of Graphics, which allowed us to create complex and aesthetically pleasing visualizations with ease.

- `knitr` [@knitr]: This package enabled us to perform data demonstration with tables.

- `styler` [@styler]: This package is helpful for styling the code.

- `arrow` [@arrow]: This package provides a convenient and efficient way to work with parquet format.

- `validate` [@validate]: This package provide useful functions for data tests.

We are grateful to the Consumer Financial Protection Bureau (CFPB) for conducting the Financial Well-Being Survey [@cfpb_fwb_survey] and making the data publicly available. Their dedication to collecting and sharing such vital information significantly contributes to research on financial well-being.

Finally, we would like to thank all those who contributed to the development and maintenance of the R programming language and its ecosystem, as well as the broader open-source community, whose efforts make such research possible.

# References